{\bf Practical Concurrency Testing} \\ \normalsize \vspace{1em} \begin{tabular}{c} {\em or: How I Learned to Stop Worrying and Love the Exponential Explosion} \end{tabular}

### 2018-12-26 thesis/education.tex
cpu-time or cpu time

The survey responses were very positive:
% corroborate this with snapshot data? -- haha, never
students reported being able both to diagnose and to verify as fixed the vast majority of Landslide's reported bugs,
compared Landslide favourably to stress testing,
and found the experience worthwhile and worth recommending.
Regarding the 100 CPU-hours question in particular,
students could approach answering it in several different ways:
do the three students who answered 20/80 think stress testing is four times as good as Landslide at finding bugs,
or that stress testing requires four times as long to reach the same point of diminishing returns?
On the other extreme, one student said they would spend all 100 CPU-hours on Landslide,
which probably speaks more to their enthusiasm than to a careful attempt to maximize expected number of bugs found
(even the author themself recognizes stress testing's advantages for certain types of bugs,
such as resource exhaustion; see \cref{sec:warpzone-verif}).
Nevertheless, the responses' overall bias to spending at least half the CPU-time on Landslide
shows clearly that the students found the experience worthwhile.

>>>>

The survey responses were very positive:
% corroborate this with snapshot data? -- haha, never
students reported being able both to diagnose and to verify as fixed the vast majority of Landslide's reported bugs,
compared Landslide favourably to stress testing,
and found the experience worthwhile and worth recommending.
Regarding the 100 CPU-hours question in particular,
students could approach answering it in several different ways:
do the three students who answered 20/80 think stress testing is four times as good as Landslide at finding bugs,
or that stress testing requires four times as long to reach the same point of diminishing returns?
On the other extreme, one student said they would spend all 100 CPU-hours on Landslide,
which probably speaks more to their enthusiasm than to a careful attempt to maximize expected number of bugs found
(even the author themself recognizes stress testing's advantages for certain types of bugs,
such as resource exhaustion; see \cref{sec:warpzone-verif}).
Nevertheless, the responses' overall bias to spending at least half the CPU time on Landslide
shows clearly that the students found the experience worthwhile.

### 2018-12-26 thesis/quicksand.tex
cpu-time or cpu time

Given unlimited CPU time, %for testing,
%one would always want to
switching to the new maximal state space whenever adding a new preemption point
\qrevisionminor{would be the quickest way to reach full verification}.
The maximal state space is guaranteed to subsume all execution sequences reachable in any subset state space,
so considering any incomplete subset of the known preemption points would be redundant work.
\Cref{alg:algorithm0} demonstrates this na\"ive approach.
It is seeded with the set of all statically-known synchronization API preemption points,
and invoked whenever a new data race candidate is found.
%
The upcoming proofs in \cref{sec:quicksand-soundness},
being concerned with the verification guarantee provided when the search may complete within the CPU budget,
are based on this simple version of Iterative Deepening.
The user may also wish to configure her testing tool to prefer this approach, at her discretion,
such as when she believes all bugs have been fixed and wants a verification as fast as possible;
\cref{sec:quicksand-impl-modes} discusses this execution mode further.

@@

Prior work has shown how to parallelize DPOR of a single state space across multiple processors \cite{parallel-dpor},
but it remains an open research problem how to extend the algorithm to ICB.
Hence, I optimistically gave all control experiments a linear speedup of 10 hours per test with 1 CPU;
i.e., assuming it could parallelize with 100\% efficiency.
To match this,
Quicksand reports both the CPU-time and wall-clock time spent during its execution.
Comparing CPU-time leads to a more fair comparison,
although Quicksand's inherent parallelism, which only a wall-clock time comparison would show,
is also a convenient benefit unto itself.
All tests ran on 12-core 3.2 GHz Xeon W3670 machines with 12GB of RAM.

@@

\Cref{fig:dowefindbugsfaster-cpu}
plots the bug-finding performance of Quicksand's three experimental trials
against the three control approaches
in a cumulative distribution of total bugs found against elapsed CPU-time.
The farthest-right point on each series indicates in how many total test cases that trial found a bug
after the 10 CPU-hour timeout.
\Cref{fig:dowefindbugsfaster-wall} shows
the same experiments, measured by wall-clock time until each bug was found instead.

>>>>

Given unlimited testing time, %for testing,
%one would always want to
switching to the new maximal state space whenever adding a new preemption point
\qrevisionminor{would be the quickest way to reach full verification}.
The maximal state space is guaranteed to subsume all execution sequences reachable in any subset state space,
so considering any incomplete subset of the known preemption points would be redundant work.
\Cref{alg:algorithm0} demonstrates this na\"ive approach.
It is seeded with the set of all statically-known synchronization API preemption points,
and invoked whenever a new data race candidate is found.
%
The upcoming proofs in \cref{sec:quicksand-soundness},
being concerned with the verification guarantee provided when the search may complete within the CPU budget,
are based on this simple version of Iterative Deepening.
The user may also wish to configure her testing tool to prefer this approach, at her discretion,
such as when she believes all bugs have been fixed and wants a verification as fast as possible;
\cref{sec:quicksand-impl-modes} discusses this execution mode further.

@@

Prior work has shown how to parallelize DPOR of a single state space across multiple processors \cite{parallel-dpor},
but it remains an open research problem how to extend the algorithm to ICB.
Hence, I optimistically gave all control experiments a linear speedup of 10 hours per test with 1 CPU;
i.e., assuming it could parallelize with 100\% efficiency.
To match this,
Quicksand reports both the CPU time and wall-clock time spent during its execution.
Comparing CPU time leads to a more fair comparison,
although Quicksand's inherent parallelism, which only a wall-clock time comparison would show,
is also a convenient benefit unto itself.
All tests ran on 12-core 3.2 GHz Xeon W3670 machines with 12GB of RAM.

@@

\Cref{fig:dowefindbugsfaster-cpu}
plots the bug-finding performance of Quicksand's three experimental trials
against the three control approaches
in a cumulative distribution of total bugs found against elapsed CPU time.
The farthest-right point on each series indicates in how many total test cases that trial found a bug
after the 10 CPU-hour timeout.
\Cref{fig:dowefindbugsfaster-wall} shows
the same experiments, measured by wall-clock time until each bug was found instead.

### 2018-12-26 thesis/tm.tex
cpu-time or cpu time

The notation {\tt testname}$(K,N)$ %will denote
denotes a test configuration of $K$ threads, each running $N$ iterations of the test logic.
All tests were run on an 8-core 2.7GHz Core i7 with 32 GB RAM.
Reported CPU-times include time spent on all state spaces
Quicksand saw fit to run,
not just the maximal or the buggy state space;
for verification tests (run with {\tt -M}),
this still includes abandoned smaller jobs that were run to saturate the set of data-race preemption points
(\Cref{chap:quicksand}).
To minimize variance in CPU-time measurements,
I ensured the test machine was not loaded beyond normal web browser use,
and ran only one instance of Landslide at a time;
for further discussion of variance see \cite{sigbovik-htm}.
The number of interleavings in each state space \revisionminor{is}, of course, deterministic and do not vary across runs.

@@

		For each, I list the CPU-time and wall-clock time elapsed,

@@

Quicksand's ability to find bugs in fewer distinct interleavings (i.e., overall smaller state spaces)
does not necessarily correlate with better performance in terms of CPU-time.
Comparing \Cref{tab:tm-buges}'s trends
to the break-even point in Quicksand's evaluation (\cref{sec:quicksand-eval}),
most of these tests are too small for its approach to pay off,
with {\tt swapbug} and {\tt avl\_insert} as its notable wins.
While plenty more wins were observed in \cref{sec:quicksand-eval},
%an upper bound also exists:
%in {\tt swapbug}(3,2),
%not even the {\em minimal} state space was completed in time.
this suggests
future MCs could prioritize state spaces using not just size estimation
but a hybrid approach also conisdering state space maximality and preemption bounds \cite{chess-icb}
to soften the trade-off
both for smaller tests and for verification.

>>>>

The notation {\tt testname}$(K,N)$ %will denote
denotes a test configuration of $K$ threads, each running $N$ iterations of the test logic.
All tests were run on an 8-core 2.7GHz Core i7 with 32 GB RAM.
Reported CPU times include time spent on all state spaces
Quicksand saw fit to run,
not just the maximal or the buggy state space;
for verification tests (run with {\tt -M}),
this still includes abandoned smaller jobs that were run to saturate the set of data-race preemption points
(\Cref{chap:quicksand}).
To minimize variance in CPU-time measurements,
I ensured the test machine was not loaded beyond normal web browser use,
and ran only one instance of Landslide at a time;
for further discussion of variance see \cite{sigbovik-htm}.
The number of interleavings in each state space \revisionminor{is}, of course, deterministic and do not vary across runs.

@@

		For each, I list the CPU time and wall-clock time elapsed,

@@

Quicksand's ability to find bugs in fewer distinct interleavings (i.e., overall smaller state spaces)
does not necessarily correlate with better performance in terms of CPU time.
Comparing \Cref{tab:tm-buges}'s trends
to the break-even point in Quicksand's evaluation (\cref{sec:quicksand-eval}),
most of these tests are too small for its approach to pay off,
with {\tt swapbug} and {\tt avl\_insert} as its notable wins.
While plenty more wins were observed in \cref{sec:quicksand-eval},
%an upper bound also exists:
%in {\tt swapbug}(3,2),
%not even the {\em minimal} state space was completed in time.
this suggests
future MCs could prioritize state spaces using not just size estimation
but a hybrid approach also conisdering state space maximality and preemption bounds \cite{chess-icb}
to soften the trade-off
both for smaller tests and for verification.
